---
title: "Central Limit Theorm"
author: "Chathura"
date: "October 18, 2016"
output: html_document
---
Central Limit Theorm may be the most important statistical concept. Briefly we can state that for any distribution with *finite mean and standard deviation*, samples taken from that population will tend towards a normal distribution around the mean of the population as sample size increases. Furthermore, as sample size increases, the variation of the sample means will decrease.

Lets see this with an simulation. suppose we have one six sided die. Drawing a sample mean playing the die and observing the outcome. Obviously we will get one value for each time we draw. That means we are taking a sample from (1,2,3,4,5,6) one at a time. Each outcome is equally likely. so if we draw 1000 times seperatly. we can plot a histogram like this.
```{r}
numoftimes = 10000
possibleSides <- 1:6
x <- as.integer(runif(numoftimes,1,7))
par(mfrow=c(1,2))
hist(x,main=paste( numoftimes," roles of a single die"),breaks=seq(1-.5,6+.5,1))
boxplot(x, horizontal=TRUE,range=1)
title("boxplot of a uniform random distribution")
```
So our population has 6 items. each one is equally likly. Mean and the variance of the population can be calculated as above.
lets introduce the random variable concept. If we  play the die, we get a value, which is number governed by a complex natural process which we can not predict. We represent the RV by:

$$X$$

If we play the die 1000 times and get the average of the values we get. it is called **Expected Value**.we write it as 
$$[X] = (1+2+3+4+5+6)/6$$.

if we draw 1000 times, we drawing a sample of size 1 from the population of 1,2,3,4,5,6. so mean of the each sample can be from 1 to 6. mean of sample means and the variance of the sample variance can be shown as below. Mean and variance of the sampling distribution are:
```{r, echo=FALSE}
mean(x)
var(x)
hist(x,main=paste( numoftimes," Sampling distribution when sample size is 1"),breaks=seq(1-.5,6+.5,1))
```
and the distribution has large variance.
if we take 2 random values from the same population of 1 to 6. now our sample can be anything from (1,1)..to...(6,6).
```{r}
sampling_means<-c()
for(i in 1:10000){
  #roll the first die
die1<-floor(runif(1,min=1,max=7))
  #roll the second die
die2<-floor(runif(1,min=1,max=7))
sample_mean = sum(die1,die2)/2
sampling_means<-c(sampling_means,sample_mean)
}
hist(sampling_means,main = "histogram of means of samples of size2")
mean(sampling_means)
var(sampling_means)
```
If we sample 3 dies. we get a more normal distribution with lesser variance and mean is very close to population mean.
```{r}
sampling_means<-c()
for(i in 1:10000){
  #roll the first die
die1<-floor(runif(1,min=1,max=7))
  #roll the second die
die2<-floor(runif(1,min=1,max=7))

die3<-floor(runif(1,min=1,max=7))

sample_mean = sum(die1,die2,die3)/3
sampling_means<-c(sampling_means,sample_mean)
}
hist(sampling_means,main = "histogram of means of samples of size3")
mean(sampling_means)
var(sampling_means)
```





